Methods for Analyzing Models of Rod-Shaped Bacteria

### 2025-07-18 paper/main.tex
move multilayer stuff to supplement

\begin{table}[H]
    \centering
    \def\arraystretch{1.3}
    \begin{tabularx}{\textwidth}{c l X}
        &\textbf{Aspect} & \textbf{Description}\\
        \toprule
        &\textbf{(C) Cellular}\\
        \midrule
        (1) & Rod-Shaped Mechanics &
            Rod-shaped bacteria are flexible rods which are able to freely move around (off-lattice
            approach) \cite{Takeuchi2005,Ursell2014,Amir2014_2}.\\
        (2) & Growth &
            Cells grow exponentially by inserting new material along the circular part of the rod
            \cite{Robert2014,Takeuchi2005}.\\
        % (3) & Differentiation &
        %     \textit{B.subtilis} is known to differentiate into matrix-producing and
        %     surfactin-producing cells \cite{vanGestel2015,Lpez2010}.\\
        (3) & Division &
            Rod-Shaped bacteria divide in the middle of the rod, thus forming two new
            agents.
            Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
        (4) & Variable Parameters &
            Parameters for individual cells are not fixed values but rather taken from a
            distribution \cite{Koutsoumanis2013}.\\
        &\textbf{(CC) Cell-Cell Interactions}\\
        \midrule
        (5) & Adhesion &
            Bacteria adhere to each other at longer distances and attach when in close contact
            \cite{Verwey1947,Trejo2013}.\\
            % The interaction between rods can be polarized \cite{Duvernoy2018}.\\
        % (7) & Friction &
        %     Friction between cells \cite{Grant2014} can be asymmetrical \cite{Doumic2020}.\\
        &\textbf{(DC) Domain-Cell Interactions}\\
        \midrule
        (6) & External Forces &
            Bacteria which rub against surfaces tend to stick to them after some
            time~\cite{vanLoosdrecht1989}.
            The extracellular gel which surrounds the bacteria exerts a force onto the cells
            \cite{Grant2014}.\\
        % (9) & Extracellular Reactions &
        %     Bacteria can take up nutrients or possible secrete/take up signalling molecules
        %     \cite{Li2025}.\\
        \bottomrule
    \end{tabularx}
    \label{table:simulation-aspects}
    \caption{TODO Insert caption}
\end{table}

@@

% --------------------------------------------------------------------------------------------------
\subsection{External Forces}
\label{subsection:mechanical-model-external-forces}
TODO
\begin{itemize}
    \item Friction force against surface. perpendicular to motion; proportional to velocity.
    \item gel pressure coming from experiment (simply downwards force $F_z$)
\end{itemize}

@@

%###################################################################################################
\section{Multilayer Predictions}
% TODO
\begin{itemize}
    \item Do we need this in this publication?
    \item Maybe this is out of scope as this is a more methodological paper.
    \item If we want to omit it, we might also opt to change the model description in the beginning
        and remove some parts of the external forces discussion.
\end{itemize}

>>>>

\begin{table}[H]
    \centering
    \def\arraystretch{1.3}
    \begin{tabularx}{\textwidth}{c l X}
        &\textbf{Aspect} & \textbf{Description}\\
        \toprule
        &\textbf{(C) Cellular}\\
        \midrule
        (1) & Rod-Shaped Mechanics &
            Rod-shaped bacteria are flexible rods which are able to freely move around (off-lattice
            approach) \cite{Takeuchi2005,Ursell2014,Amir2014_2}.\\
        (2) & Growth &
            Cells grow exponentially by inserting new material along the circular part of the rod
            \cite{Robert2014,Takeuchi2005}.\\
        % (3) & Differentiation &
        %     \textit{B.subtilis} is known to differentiate into matrix-producing and
        %     surfactin-producing cells \cite{vanGestel2015,Lpez2010}.\\
        (3) & Division &
            Rod-Shaped bacteria divide in the middle of the rod, thus forming two new
            agents.\\
            % Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
        (4) & Variable Parameters &
            Parameters for individual cells are not fixed values but rather taken from a
            distribution \cite{Koutsoumanis2013}.\\
        &\textbf{(CC) Cell-Cell Interactions}\\
        \midrule
        (5) & Adhesion &
            Bacteria adhere to each other at longer distances and attach when in close contact
            \cite{Verwey1947,Trejo2013}.\\
            % The interaction between rods can be polarized \cite{Duvernoy2018}.\\
        % (7) & Friction &
        %     Friction between cells \cite{Grant2014} can be asymmetrical \cite{Doumic2020}.\\
        % &\textbf{(DC) Domain-Cell Interactions}\\
        % \midrule
        % (6) & External Forces &
        %     Bacteria which rub against surfaces tend to stick to them after some
        %     time~\cite{vanLoosdrecht1989}.
        %     The extracellular gel which surrounds the bacteria exerts a force onto the cells
        %     \cite{Grant2014}.\\
        % (9) & Extracellular Reactions &
        %     Bacteria can take up nutrients or possible secrete/take up signalling molecules
        %     \cite{Li2025}.\\
        \bottomrule
    \end{tabularx}
    \label{table:simulation-aspects}
    \caption{TODO Insert caption}
\end{table}

@@

@@

%###################################################################################################
\section{Multilayer Predictions}
% TODO

\begin{itemize}
    \item Friction force against surface. perpendicular to motion; proportional to velocity.
    \item gel pressure coming from experiment (simply downwards force $F_z$)
\end{itemize}

\begin{itemize}
    \item Do we need this in this publication?
    \item Maybe this is out of scope as this is a more methodological paper.
    \item If we want to omit it, we might also opt to change the model description in the beginning
        and remove some parts of the external forces discussion.
\end{itemize}
